---
title: "r-py-data-types"
format: html
editor: visual
---

```{r}
dogs <- c("teddy","khora","waffle", "banjo")
class(dogs)
typeof(dogs)
```
#numeric vector
```{r}
weights <- c(50,55,25,35)
typeof(weights)
class(weights)
```
#integer vector
```{r}
dog_age <- c(5L,6L,1L,7L)
typeof(dog_age)
class(dog_age)
is.numeric(dog_age)

```
#check with a logical
```{r}

is.numeric(dog_age)
```
#combining classes
```{r}
dog_info <- c("teddy", 50, 5L)
dog_info
class(dog_info)
```

```{r}
typeof(dog_info)
class(dog_info)
is.character(dog_info)
```
```{r}
#named elements
dog_food <- c(teddy = "purina", 
              khora = "alpo", 
              waffle = "feast", 
              banjo = "blue diamond")
dog_food
```


```{r}
class(dog_food)
```

```{r}
typeof(dog_food)
```


```{r}
#indexing and getting vectors
dog_food[2]
```

```{r}
dog_food["khora"]
dog_food[1:3]
```

#Create vectors of car colors observed

```{r}
cars <- c("red", "orange", "white", "blue", "green", "silver", "black")
#access the 5th element

cars[5]
```

# Access elements 2 - 4

```{r}
cars[2:4]
```


#warm up to for loops
```{r}
i <- 4
cars[i]

```
```{r}
i <- seq(1:3)
cars[i]
```
#updating the elements of a vector directly (mutable)

```{r}
cars[3] <- "BURRITOS"
cars
```
#Matrices


```{r}
fish_size <- matrix(c(0.8, 1.2,0.4,0.9), ncol = 2, nrow = 2, byrow = FALSE)
fish_size
```

```{r}
typeof(fish_size)
```
```{r}
class(fish_size) #returns matrix / array
```

```{r}
dog_walk <- matrix(c("teddy", 5, "khora", 10), ncol = 2, nrow = 2, byrow = FALSE)
dog_walk
```
```{r}
class(dog_walk)
typeof(dog_walk)
```

# Accessing pieces of matrices

```{r}
 whale_travel <- matrix(data = c(31.8,1348, 46.9, 1587), nrow = 2, ncol = 2, byrow = TRUE)
#take a look

whale_travel
```

#Accessing pieces of matrices

```{r}
 whale_travel <- matrix(data = c(31.8,1348, 46.9, 1587), nrow = 2, ncol = 2, byrow = TRUE)

#take a look
whale_travel
```

#Access value 1348

```{r}
whale_travel[1,2] # row1, col 2
```
```{r}
whale_travel[2, ]
```


```{r}
whale_travel[,1]
```

```{r}
#what happens if we give a matrix with one element? THats the position in the matric as if populated by column
whale_travel[3]
```
# Lists

```{r}
urchins <- list("blue", c(1,2,3), c("a cat", "a dog"), 5L)
urchins
```
#Accessing pieces ofd a list

```{r}
urchins[[2]]
```
#compare that to:
```{r}
urchins[2]
```
#Naming list items?

```{r}
tacos <- list(topping = c("onion", "cilantro", "guacamole"), filling = c("beans", "meat", "veggie"), price = c(6.75, 8.25, 9.50))

#the whole thing
tacos
```
```{r}
#just get one piece of it:
tacos[[2]]


#or the same thing
tacos$filling
```

```{r}
# Create a dataframe
fruit <- data.frame(type = c("apple", "banana", "peach"), mass = c(130, 195, 150))

#look at it
fruit
```

```{r}
fruit[1,2]
```

```{r}
fruit[3,1]
```

```{r}
#changing the value of column 1, row 2
fruit[2,1] <- "pineapple"
fruit
```

# Making & Indexing data in Python
#in diff file (python)
```

















